>

### 2019-05-03 paper/spinifex-paper.tex
expanded rotation matrix and it's explaination.

Rotations are defined by the axes they pivot on, \(\theta\) then is the
rotation around the \(Z\) axis. While \(\phi\) will be the rotation
around, \(U\), a new axis on the \(XY\) plane, orthogonal to \(\phi\).
Given these axes, the rotation matrix can be derived by the following.

@@

\begin{itemize}
  \item[$c_\theta$] be the cosine of $\theta$
  \item[$c_\phi$]   be the cosine of $\phi_i$
  \item[$s_\theta$] be the sine of   $\theta$
  \item[$s_\phi$]   be the sine of   $\phi_i$
\end{itemize}

\begin{align*}
  \textbf{B}_k &= (c_\theta, s_\theta, 0) \\
  \vec{u} &= (s_\theta, -c_\theta, 0) \\ &=
  \begin{bmatrix}
     0 & 0 & -c_\theta \\
     0 & 0 & -s_\theta \\
     c_\theta & s_\theta & 0 \\
   \end{bmatrix} \\
  \text{Apply Rodrigues' rotation formula:} \\
    \textbf{R}_{[3,~3]} 
    &= \textbf{I}_3 + s_\phi\*\vec{u} + (1-c_\phi)\*\vec{u}^2 \\
    &= 
    \begin{bmatrix}
      c_\theta^2 c_\phi + s_\theta^2 &
      -c_\theta s_\theta (1 - c_\phi) &
      -c_\theta s_\phi \\
      -c_\theta s_\theta (1 - c_\phi) &
      s_\theta^2 c_\phi + c_\theta^2 &
      -s_\theta s_\phi \\
      c_\theta s_\phi &
      s_\theta s_\phi &
      c_\phi
    \end{bmatrix}
\end{align*}

Post-multiply the manipulation space by the pre-defined rotation matrix
producing \textbf{RM}, the rotated manip space. This is one frame of
basis values, repeat this process for each value in the sequence
\(\phi_i\) for the complete animation.

>>>>

Rotations are defined by the axes they pivot on, \(\theta\) then is the
rotation around the \(Z\) axis. While \(\phi\) will be the rotation
around, \(\textbf{U}\), a new axis on the \(XY\) plane, orthogonal to
\(\phi\). Given these axes, the SO(3) rotation matrix can be derived by
the following.

@@

\begin{itemize}
  \item[$c_\theta$] be the cosine of $\theta$
  \item[$c_\phi$]   be the cosine of $\phi_i$
  \item[$s_\theta$] be the sine of   $\theta$
  \item[$s_\phi$]   be the sine of   $\phi_i$
  \item[$\textbf{U}$] be the axis of rotation for $\phi$.
\end{itemize}

\begin{align*}
  \textbf{U} &= (u_x, u_y, u_z) =
  (s_\theta, -c_\theta, 0) \\ 
  &=
  \begin{bmatrix}
  0 & -u_z & u_y  \\
  u_z & 0 & -u_x \\
  -u_y & u_x & 0 \\
  \end{bmatrix} =
  \begin{bmatrix}
    0 & 0 & -c_\theta \\
    0 & 0 & -s_\theta \\
    c_\theta & s_\theta & 0 \\
  \end{bmatrix} \\
  \end{align*} Apply Rodrigues' rotation formula: \begin{align*}
    \textbf{R}_{[3,~3]} 
    &= \textbf{I}_3 + s_\phi\*\textbf{U} + (1-c_\phi)\*\textbf{U}^2 \\
        &=
    \begin{bmatrix}
      1&0&0 \\ 0&1&0 \\ 0&0&1 \\
    \end{bmatrix} +
    \begin{bmatrix}
      0 & 0 & c_\theta s_\phi \\
      0 & 0 & s_\theta s_\phi \\
      -c_\theta s_\phi & -s_\theta s_\phi & 0 \\
    \end{bmatrix} +
    \begin{bmatrix}
      -c_\theta (1-c_\phi) & s^2_\theta (1-c_\phi) & 0 \\
      -c_\theta s_\theta (1-c_\phi) & -s^2_\theta (1-c_\phi) & 0 \\
      0 & 0 & c_\phi-1 \\
    \end{bmatrix} \\
    &= 
    \begin{bmatrix}
      c_\theta^2 c_\phi + s_\theta^2 &
      -c_\theta s_\theta (1 - c_\phi) &
      -c_\theta s_\phi \\
      -c_\theta s_\theta (1 - c_\phi) &
      s_\theta^2 c_\phi + c_\theta^2 &
      -s_\theta s_\phi \\
      c_\theta s_\phi &
      s_\theta s_\phi &
      c_\phi
    \end{bmatrix} \\
\end{align*}

Defining the axis of rotation for \(\phi\) and the application of
Rogrigues' rotation formula results in the rotation matrix, the rotation
group SO(3). Continue with the rotation of the manipulation space,
post-multiply the space by the above rotation matrix producing
\textbf{RM}, the rotated manip space. This is basis values, the linear
combination of variables, for one frame of the animation, repeat this
process for each value in the sequence \(\phi_i\) for the complete
animation.

### 2019-05-03 paper/spinifex_paper.Rmd
expanded rotation matrix and it's explaination.

<!-- define the rotation matrix -->
Rotations are defined by the axes they pivot on, $\theta$ then is the rotation around the $Z$ axis. While $\phi$ will be the rotation around, $U$, a new axis on the $XY$ plane, orthogonal to $\phi$. Given these axes, the rotation matrix can be derived by the following.

Let:
\begin{itemize}
  \item[$c_\theta$] be the cosine of $\theta$
  \item[$c_\phi$]   be the cosine of $\phi_i$
  \item[$s_\theta$] be the sine of   $\theta$
  \item[$s_\phi$]   be the sine of   $\phi_i$
\end{itemize}

\begin{align*}
  \textbf{B}_k &= (c_\theta, s_\theta, 0) \\
  \vec{u} &= (s_\theta, -c_\theta, 0) \\ &=
  \begin{bmatrix}
     0 & 0 & -c_\theta \\
     0 & 0 & -s_\theta \\
     c_\theta & s_\theta & 0 \\
   \end{bmatrix} \\
  \text{Apply Rodrigues' rotation formula:} \\
    \textbf{R}_{[3,~3]} 
    &= \textbf{I}_3 + s_\phi\*\vec{u} + (1-c_\phi)\*\vec{u}^2 \\
    &= 
    \begin{bmatrix}
      c_\theta^2 c_\phi + s_\theta^2 &
      -c_\theta s_\theta (1 - c_\phi) &
      -c_\theta s_\phi \\
      -c_\theta s_\theta (1 - c_\phi) &
      s_\theta^2 c_\phi + c_\theta^2 &
      -s_\theta s_\phi \\
      c_\theta s_\phi &
      s_\theta s_\phi &
      c_\phi
    \end{bmatrix}
\end{align*}

<!-- rotate manip space -->
Post-multiply the manipulation space by the pre-defined rotation matrix producing **RM**, the rotated manip space. This is one frame of basis values, repeat this process for each value in the sequence $\phi_i$ for the complete animation.

>>>>

<!-- define the rotation matrix -->
Rotations are defined by the axes they pivot on, $\theta$ then is the rotation around the $Z$ axis. While $\phi$ will be the rotation around, $\textbf{U}$, a new axis on the $XY$ plane, orthogonal to $\phi$. Given these axes, the SO(3) rotation matrix can be derived by the following.

Let:
\begin{itemize}
  \item[$c_\theta$] be the cosine of $\theta$
  \item[$c_\phi$]   be the cosine of $\phi_i$
  \item[$s_\theta$] be the sine of   $\theta$
  \item[$s_\phi$]   be the sine of   $\phi_i$
  \item[$\textbf{U}$] be the axis of rotation for $\phi$.
\end{itemize}

\begin{align*}
  \textbf{U} &= (u_x, u_y, u_z) =
  (s_\theta, -c_\theta, 0) \\ 
  &=
  \begin{bmatrix}
  0 & -u_z & u_y  \\
  u_z & 0 & -u_x \\
  -u_y & u_x & 0 \\
  \end{bmatrix} =
  \begin{bmatrix}
    0 & 0 & -c_\theta \\
    0 & 0 & -s_\theta \\
    c_\theta & s_\theta & 0 \\
  \end{bmatrix} \\
  \end{align*}
Apply Rodrigues' rotation formula:
  \begin{align*}
    \textbf{R}_{[3,~3]} 
    &= \textbf{I}_3 + s_\phi\*\textbf{U} + (1-c_\phi)\*\textbf{U}^2 \\
        &=
    \begin{bmatrix}
      1&0&0 \\ 0&1&0 \\ 0&0&1 \\
    \end{bmatrix} +
    \begin{bmatrix}
      0 & 0 & c_\theta s_\phi \\
      0 & 0 & s_\theta s_\phi \\
      -c_\theta s_\phi & -s_\theta s_\phi & 0 \\
    \end{bmatrix} +
    \begin{bmatrix}
      -c_\theta (1-c_\phi) & s^2_\theta (1-c_\phi) & 0 \\
      -c_\theta s_\theta (1-c_\phi) & -s^2_\theta (1-c_\phi) & 0 \\
      0 & 0 & c_\phi-1 \\
    \end{bmatrix} \\
    &= 
    \begin{bmatrix}
      c_\theta^2 c_\phi + s_\theta^2 &
      -c_\theta s_\theta (1 - c_\phi) &
      -c_\theta s_\phi \\
      -c_\theta s_\theta (1 - c_\phi) &
      s_\theta^2 c_\phi + c_\theta^2 &
      -s_\theta s_\phi \\
      c_\theta s_\phi &
      s_\theta s_\phi &
      c_\phi
    \end{bmatrix} \\
\end{align*}
<!-- The term $(1-c_\phi)$ is used by convention, but $2sin^2(\phi/2)$ is more computationally robust. -->

<!-- rotate manip space -->
Defining the axis of rotation for $\phi$ and the application of Rogrigues' rotation formula results in the rotation matrix, the rotation group SO(3). Continue with the rotation of the manipulation space, post-multiply the space by the above rotation matrix producing **RM**, the rotated manip space. This is basis values, the linear combination of variables, for one frame of the animation, repeat this process for each value in the sequence $\phi_i$ for the complete animation.
